### 2018-07-31 paper/paper.tex
Paper review and update figures

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

% tablespacing
\renewcommand{\arraystretch}{1.5}
\setlength\tabcolsep{0.2cm}

\input{solidity-highlighting.tex}

\lstset{basicstyle=\small\ttfamily}

\begin{document}
%
\title{Untrusted B2B Collaboration using the Blockchain}
%
%
\author{Jonas Beyer \and Philipp Otto \and S\"oren Tietb\"ohl}
%
\authorrunning{J. Beyer et al.}
%
\institute{
	Hasso-Plattner-Institut, Potsdam, Germany\\
	\email{\{jonas.beyer,philipp.otto2,soeren.tietboehl\}@student.hpi.de}}
%
\maketitle % typeset the header of the contribution
%
\begin{abstract}
	Lack of trust between communication partners is one of the major issues companies face when implementing business processes in a cross-organizational setting.
	Blockchain is an emerging technology that enables decentralization and distribution of state without the need for a trusted central instance.
	As a result, it is possible to form consensus in the communication between otherwise untrusting parties.
	In this paper we introduce Armadillo, an open-source prototype implementation for realizing business process choreographies using the Ethereum blockchain.
	We present a smart contract system for inter-process communication, as well as a user interface for connecting business processes and smart contracts.
	Our solution describes a basis for untrusted messaging via the blockchain without a central controlling party.

	\keywords{Business Processes \and Choreographies \and Blockchain \and Smart Contracts}
\end{abstract}
%
\section{Introduction}
% BPM
Business Process Management (BPM) allows organizations to identify, structure and understand the tasks and processes they employ to achieve their business goals.
%Process Models
In BPM, process models are used to document how these goals can be reached, visually describing the participants, activity flow and data objects involved.

% What is BPMN?
Business Process Model and Notation (BPMN) is a notation standard for such process models~\cite{omg2014bpmn}.
It provides building blocks for the modeling of not only internal procedures, but also the message exchange needed for successful communication with external partners.

% Trust
The basis for inter-process communication is that all involved parties trust each other.
Agreements and contracts can help clarify the requirements and expectations between partners.
As soon as multiple parties are involved in the handling of a process, though, it becomes difficult to track faults and misunderstandings in the communication.

% What is Blockchain?
With the help of the emerging Blockchain technology, these problems can be resolved.
A Blockchain network is a network of computation nodes sharing a decentralized information storage structure, a \emph{ledger}.
Inside this ledger, information is stored in blocks, which are chained together via cryptograhic means.
This allows nodes in the network to agree on a public common state of information without trusting each other.
Blockchain technology is commonly used to realize virtual currencies, also called cryptocurrencies.
Ethereum~\cite{wood2014ethereum} is one implementation of the Blockchain concept which also realizes a feature called smart contracts.
Smart contracts are small programs with limited processing power that are executed by the nodes in the Blockchain network.

%What problems are there with BPMN that can be solved with Blockchain?
Blockchain features can be used to facilitate trusted communication between untrusted parties in Business Process Management.
% What approaches exist currently (Caterpillar~\cite{lopez2017caterpillar}, Untrusted paper~\cite{weber2016untrusted})
Approaches exist that aim to bring BPM functionality to the Ethereum Blockchain utilizing smart contracts.
Weber et al.~\cite{weber2016untrusted} realize process orchestrations on the Blockchain using a smart contract as the mediator for messages and execution state.
% What is our general Idea?
We expand on the work of Weber et al. and implement process choreographies, where instead of one central mediator, each party provides their own smart contracts.
Communication then happens directly between the partner's contracts instead of passing through a central instance.

% Structure
The rest of this paper is structured as follows: In Section \ref{preliminary} we will discuss concepts and technologies required for our work, and introduce our use case.
In Section \ref{solutionapproach} we will present the solution requirements and concepts.
Section \ref{implementation} describes our contract and tool implementations.
In Section \ref{evaluation} we discuss our solution.
Finally, Section \ref{futurework} will present future work and Section \ref{conclusion} will conclude the paper.

\section{Preliminary} \label{preliminary}

\subsection{Business Process Collaboration}

When businesses work together to reach a common goal, their processes also need to comply with each other.
Such business-to-business collaborations can also be modeled in BPMN.
Collaboration models include the businesses' \emph{public processes} (or \emph{behavioral interfaces}~\cite{Weske2012}), i.e. the BPMN tasks, events and message flows describing the message exchanges between participants.
While not directly part of the interaction, they can also depict \emph{private processes}, i.e. tasks and control flows that are internally executed before or after sending or receiving a message.
An example collaboration model can be seen in Figure \ref{fig:collaboration}.

Process interoperability needs to be ensured when it comes to executing a collaboration model.
To this end, there are two different implementation approaches:
One option is to treat the collaboration like a local, internal process, employing a \emph{process orchestration}.
In a process orchestration, the execution is controlled by a central authority.
Adopting the role of a mediator or manager, this central authority dictates and guides the execution.
Thus, the execution has a well-defined order of tasks and a clear execution state.

Another option is to have the participants act more independently.
In \cite{Weske2012}, Weske describes \emph{process choreographies}, which define and ensure a set of collaboration rules to ensure interoperabiltiy.
A choreography clarifies the dependencies between the participant's processes and the messages they need to exchange for a successful collaboration.
Communication and execution is not handled by a central instance, but rather by each participant on their own.
Messages are passed asynchronously without a controlling channel.
As a result, there is no clear execution state, because each participant drives their own execution.

\subsection{Blockchain}

A blockchain is a distributed ledger that allows users to find a consensus in decentralized applications.
Information is stored in a linked list of blocks and shared between all participants, i.e. all nodes in a decentralized computer network.
By only allowing new blocks to be appended to the list, a non-rewritable history is created, that at the same time represents the "truth", the common state all nodes in the network agree on.
There are multiple approaches of determining who is allowed to append a new block, altering the state.
The most common one, currently employed by Bitcoin~\cite{nakamoto2008bitcoin} and Ethereum~\cite{wood2014ethereum}, is called \emph{proof of work}:
Each new block has to contain a signature that verifies the block order.
The signature is created using data from the preceding block and is by design mathematically hard to calculate, but easy to verify.
On the one hand, this results in a race between so called \emph{mining} nodes who try and calculate such a signature, as the first valid signature verifies the new blocks position and stores its data on the ledger.
Mining nodes get rewarded for finding such signatures, creating an incentive for continuous appending of new blocks and keeping the blockchain alive.
On the other hand, other nodes can easily verify the signature and come to a consensus regarding the latest state.

Blockchain technology is mainly used for cryptocurrencies, most notably Bitcoin~\cite{nakamoto2008bitcoin}.
In more recent developments however, it is a foundation for a number of other decentralized applications.
Ethereum~\cite{wood2014ethereum} is a platform that uses a blockchain as the underlying system, and additionally allows users to write and execute code in the form of \emph{smart contracts}.
Smart contracts are Turing complete code artifacts that can be deployed on the blockchain and are accessible by accounts in the Ethereum platform.
When called by a user account or another contract, they can perform calculations, send or receive money, or store information on the blockchain ledger either directly or in the form of \emph{events}.
Events are publicly visible and can be watched by off-chain applications, allowing external programs to react to blockchain state changes.
This would otherwise not be possible, since smart contracts can not communicate with any external APIs.

When calling functions on a smart contract, users have to supply \emph{gas}, a unit that powers smart contract computations.
Contracts deplete gas for each operation.
This ensures a) that contracts are not running for an unlimited amount of time, avoiding a dead lock of the platform, and b) that computations are kept small. This is important, because all operations have to be performed by each node in the network.

Smart contracts are executed in the \emph{Ethereum virtual machine (EVM)}.
As such, they contain byte code that is interpreted by the EVM.
When developing smart contracts, there are multiple languages that compile to EVM bytecode, the most prominent one currently being \emph{Solidity}\footnote{\url{https://github.com/ethereum/solidity}}, which we also chose to implement our protoype.
Xu et al. present patterns and examples for Solidity programming in~\cite{xu2018pattern}, which we used as an orientation during development.

\subsection{Prior Work on Business Processes on the Blockchain}
In~\cite{weber2016untrusted}, a first approach of monitoring and coordinating business process interactions across organizations on the blockchain was presented.
Using a smart contract as a central \emph{monitor} or \emph{mediator} of the communication, a form of inter-process orchestration is implemented.
The mediator is also responsible for role management, i.e. participants can register users to certain roles with the mediator, who verifies these roles at runtime.
The architecture features \emph{factory contracts}, which are responsible for deploying instance contracts for each process instance.
For integrating blockchain and BPM applications, the concept of \emph{triggers} is introduced: Applications that watch events on the blockchain and trigger external API calls.

We built our implementation based on these concepts.
However, our solution differs from~\cite{weber2016untrusted} in the way we realize the execution of collaboration models.
As stated before, Weber et al. employ the central mediator contract, managing execution steps and internal state of the collaboration, achieving the orchestration of a business process collaboration.
Our goal on the other hand was to implement the collaboration as a choreography, where each participant is represented by their own contracts, and is able to verify the integrity of their interaction partners.
Thus, a central mediator is no longer needed, also resolving the question of who is responsible for the development, deployment and maintenance of the mediator contract.

In order to achieve BPMN-to-Solidity compilation,~\cite{weber2016untrusted} also describes translation rules to transform BPMN elements to Solidity code.
Further work~\cite{banuelos2017optimized} optimizes the translation process, in turn reducing Ethereum execution costs by reducing required space and operations.

Caterpillar~\cite{lopez2017caterpillar} is an open source implementation of a process engine using smart contracts.
Building on the BPMN-to-Solidity compilation from~\cite{weber2016untrusted}, it integrates translation, deployment and execution of BPMN models directly on the Ethereum Blockchain.
Processes can be started and executed via a user interface, which encapsulates the calls to the generated Solidity code in the background.

Our approach is different in that we do not try to use the Ethereum platform as a process engine, but rather use it only for message passing, and integrate it with an off-chain process engine.
Caterpillar also currently does not support the execution of process interactions, which we aim to achieve.

\subsection{Use Case}

Since we build on a lot of the work in~\cite{weber2016untrusted}, we decided to also continue with the example collaboration\footnote{A supply chain scenario reported originally in~\cite{fdhila2015collaborative}, which was simplified in~\cite{weber2016untrusted}.} used there.
Because we implemented contracts for each participant, we needed to expand the collapsed pool for the \emph{Bulk Buyer}, and add tasks and events for the corresponding public process.
Additionally, we added a new lane to the \emph{Manufacturer's} process for demonstration of the role management we implemented.
The result is depicted in Figure \ref{fig:collaboration}.

% When we present the implementation, the tasks that are modeled as BPMN message sends in the collaboration model will instead be modeled as \emph{BPMN service tasks} in the separate processes.
% This decision is based on the notion that in our implementation, all sends are a call to the Ethereum platform, which acts as the external service and takes over the actual messaging.
% This implementation detail is not relevant in the collaboration view, though, where we are concerned with the interactions.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{fig/collaboration.eps}
	\caption{Supply chain scenario from~\cite{weber2016untrusted} (modified).}
	\label{fig:collaboration}
\end{figure}

\section{Solution Approach} \label{solutionapproach}

\subsection{Requirements}

Before presenting the system, we want to give an overview of the functional and non-functional requirements it has to fulfill.

The systems purpose is to facilitate trusted communication and collaboration between untrusting partners.
Therefore, it must be possible to send messages between partners.
To ensure security and trust in the system, both partners need a way to verify the authenticity of a message.
Verification is only possible if the system keeps track of users and permissions, i.e. if it implements a form of role-based access control with a user-to-role and role-to-permission mapping.
User and permission data has to be accessible at runtime, so that the communicating partners can use it to check the authenticity of messages.
It should also be possible to modify this data at runtime, to accomodate for changes in the company structure.

In order to make this system feasible in a production environment, we also require it to be reusable, measurable and cheap.
This implies keeping the implementation small and modular, and keeping it general enough to be used with the processes of different participants.
Measurability is required for evaluation and monitoring.
Since all computation done by smart contracts comes with gas costs, we want to keep the cost of our implementation as low as possible.

\subsection{System Overview}
\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{fig/system_diagram.pdf}
	\caption{An architecture diagram showing the interactions within the system.}
	\label{fig:system_diagram}
\end{figure}

Every organization provides an interface to their public process in an \emph{instance contract}.
Additionally, \emph{factory contracts} manage these instance contracts and \emph{access contracts} are used for user rights management.
% The contract structure will be described in detail in the following sections.
We assume that all participating organizations use a process engine to execute their private processes.

When a throw event is reached within a process engine, the process engine calls the \emph{throw event function} on the company's instance contract, which in turn automatically calls the corresponding \emph{catch event function} on the instance contract of the receiver.
Thus, all communication between organizations is recorded on the blockchain.

Each organization deploys a \emph{trigger} application that continuously listens to catch events (as described in~\cite{weber2016untrusted}).
If a catch event is observed, this information is relayed to the process engine to handle the event.
The usage of the blockchain to monitor the choreography is encapsulated in the process engine and is therefore hidden from the users.

\subsection{Contract Structure}

To realize the system, we use three kinds of smart contracts.
Firstly, each organization deploys one \emph{instance contract} per process instance.
It represents the public process of the organization and is called during process execution whenever a catch or throw event is reached in the execution of a process instance.
In the case of catch events, an Ethereum event is logged to the blockchain to document the process event.

Secondly, each organization deploys one \emph{factory contract} per process.
It is responsible for creating instance contracts and saving a mapping from instance IDs to the corresponding contracts.

Thirdly, each organization deploys \emph{access contracts} to manage which users can trigger events on the instance contracts.
Either there is one access contract per process or access contracts are shared over multiple processes depending on the use case.

\subsection{Per-Company Setup}

Each organization needs to perform certain setup steps before the process execution.
The factory and access contracts corresponding to the process need to be deployed.
Additionally, the user-to-role and and role-to-permission mappings have to be updated in the access contract to allow verification of users during the process execution.
Lastly, the addresses of the factory contracts of all communication partners need to be linked in the factory contract.

\subsection{Interaction Setup}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{fig/instance_creation.pdf}
	\caption{Example sequence diagram describing the creation of an instance contract of the bulk buyer, cf. Figure \ref{fig:collaboration}.}
	\label{fig:instance_creation}
\end{figure}

In order to create a new process instance, the party initiating the choreography has to explicitly call \texttt{createInstance} on their factory contract.
The factory contract in turn may need to initialize other instance contracts by calling their corresponding factory contracts.
When all partner instance, factory and access contracts are known, the instance contract can be created, its address is returned to the user and an event is emitted to the blockchain.
This process is visualized in Figure \ref{fig:instance_creation}.

The recursive initialization ensures that all instance contracts that are part of the process instance will be created by one function call.
The organizations which do not initiate the choreography can detect the newly created instance contracts by watching their factory contracts using a trigger.

\subsection{Sending Messages}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{fig/event_sending.pdf}
	\caption{A sequence diagram describing the process of sending a message to a collaboration partner, in this instance the message \emph{request details} that is sent by the special carrier to the supplier in Figure \ref{fig:collaboration}.}
	\label{fig:event_sending}
\end{figure}

%TODO correctly show that the address is passed instead of using msg.sender
When a throw event is triggered in the process engine of organization $A$, the process engine of $A$ calls the corresponding \emph{throw event function} on the instance contract of organization $A$.
In the instance contract of $A$ an authorization is performed by looking up the sender address in the access contract of $A$.
If this check is successful, the \emph{catch event function} is called on the partner instance contract of $B$.
In this instance contract there is another authorization against the access contract of $A$.
This ensures that both sides can be confident that the message was sent by a person who is authorized by organization $A$.
If both checks are successful, an event is emitted to the blockchain.
This event will be detected by the trigger of $B$ and passed to their process engine.
Thus, a message was sent from $A$ to $B$ using the blockchain as a ledger and an audit trail was created.
An example can be seen in the sequence diagram in Figure \ref{fig:event_sending}.
% TODO explain how payload could look like

\section{Implementation} \label{implementation}

Given the architecture, we will now discuss the implementation of the contracts used in the system.
Afterwards, we will shortly describe the tools we built for interacting with the system.

\subsection{Contracts}

In the following, we will formally describe our contract implementations.
The specific implemented contracts can be found in our repository in the folder\\\url{https://github.com/p-otto/Armadillo/tree/master/armadillo/resources}.

\subsubsection{Process Contracts}

For taking part in the process collaboration, each company needs to deploy a factory contract that is responsible for creating instance contracts.
To this end, the factory contract source code needs to contain the definition of the company's instance contract.
This closely follows the factory contract pattern used in~\cite{weber2016untrusted}, which is also described as pattern 14 in~\cite{xu2018pattern}.

In our implementation, a companys factory contract also knows the address of the companys access contract for convenience.
Additionally, it stores the addresses of created instances, mapped by a \emph{choreography instance ID}, i.e. an ID that is shared between all choreography participants for each instance of the shared choreography.

To allow interaction between companies via the blockchain, two participants need to exchange the addresses of their factory contracts when there is a direct interaction between them.

A factory contract thus has the resulting interface:
\begin{itemize}
	\item A \texttt{constructor} where the address of the companys access contract is passed
	\item Setter methods for the factory contract addresses of interaction partners
	\item A \texttt{createInstance} function
	\item A \texttt{getAccessAddress} function
	\item A \texttt{isInstance} function used to verify whether an address points to an instance contract created by this factory (for a given choreography ID)
\end{itemize}

Regarding storage, private fields are necessary for holding the factory contract addresses, the access contract address as well as the instance address mapping.
To avoid communication errors at runtime, it should not be possible to call \texttt{createInstance} before all required factory addresses are set.
Finally, there is one minor implementation difference depending on whether a participant is the \emph{choreography initiator} (the participant to manually start the choreography instance by sending a request):
The initiator needs to generate the choreography instance ID inside of its \texttt{createInstance} function, and pass it to all its interaction partners.
Every other participant needs to have a parameter for the ID in their \texttt{createInstance} function.

Communication between smart contracts happens in the form of function calls.
To allow the factory and instance contracts to call functions on their interaction partners contracts, the partners function interface needs to be known, i.e. it needs to be declared in the factory contract source code.
%- the access contract (same for every participant)
%- remote factory contracts (if receiving messages from remote, for verification)
%- remote instance contracts (if sending messages to remote instance)

The instance contract definition is also part of the factory contract source code.
The factory contract creates an instance by calling its constructor and passing all addresses that are required for the inter-process communication .
Instance contracts provide functions for each sending action, i.e. send tasks or events in the companys process, which can be called by the process engine.
For each receiving action (task or event), a function needs to be present that can be called from other instance contracts.
When company $A$ triggers a sending function on their instance contract, it in turn calls a corresponding receive function on the instance contract of company $B$.

Instance contracts need to store:
\begin{itemize}
	\item The choreography instance ID for verifying incoming messages
	\item The addresses of the factory and access contracts of partners for verifying incoming messages
	\item The addresses of the instance contracts of partners for sending messages
	\item The address of the companys own access contract, for verifying local function calls
	\item The address of the companys own factory contract, for returning any remaining funds upon \texttt{selfdestruct}
\end{itemize}

Verification is implemented on sender side and on receiver side.
Each send function has a guard checking the caller in the companys own access contract.
Each receive function has a guard checking a) that the caller is a contract instance that is part of the choreography instance (via the partner factory contract), and b) checking the account behind the request in the partners access contract.
Guarding functions against unwanted access was described in pattern 13, Embedded Permission, in~\cite{xu2018pattern}.

Having achieved a traceable communication on the blockchain using contract function calls, we still want to access messages in off-chain applications, e.g. triggering process events in a process engine on message receive.
To this end, smart contracts need to implement blockchain events that are emitted to the blockchain ledger whenever a receive function is executed.
These events can then be watched, and reacted upon, by off-chain components.

In our current implementation, the end of a process instance is marked by the instance contract self-destructing.
At the end of the execution of the last send or receive function in the process, the solidity builtin \texttt{selfdestruct} is called.
By passing the factory contract address, potential remaining funds are transferred to the factory contract.

\subsubsection{Access Contract}

Each company deploys an access contract that holds information about the roles and permissions of the companys employees.
Having permission data encapsuled in a separate access contract follows pattern 12, the Data Contract pattern, from~\cite{xu2018pattern}.
It allows decoupling the roles and permissions from the actual contract instances, allowing for a simple modification at runtime.
Access contracts can be deployed per process, or even once per company as a single source of truth for all processes.

To allow the administration of user permissions, access contracts provide the following interface:
\begin{itemize}
	\item \texttt{assignRoleToUser(userAddress, role)}
	\item \texttt{removeRoleFromUser(userAddress, role)}
	\item \texttt{assignTaskToRole(task, role)}
	\item \texttt{removeTaskFromRole(task, role)}
\end{itemize}
The access to these functions should be guarded, allowing only a specific administrator address to call them.

Verifying a users permission to execute a certain task can be achieved by calling \texttt{isAuthorized(userAddress, task)}.
\newline

Given this interface, there are multiple possible implementation approaches for access contracts.
We decided on a simple mapping-based implementation:
\begin{lstlisting}[
  caption=Mappings in our access contract implementation,
  language=Solidity
]
pragma solidity ^0.4.23;

contract Access {
    mapping(string => string) taskToRole;
    mapping(string => uint) roleToIndex;
    mapping(address => uint) addressToAccessBitmask;
    ...
}
\end{lstlisting}
The first mapping stores the tasks that can be executed by a specific role (cf. pools and lanes, and their contained tasks, in BPMN diagrams).
The second mapping assigns each role an index in the access bitmask, i.e. if the bit at the given index is set, the user holds the corresponding role.
The third mapping assigns such an access bitmask to each user.

\subsection{Tools}

In the following, we will briefly present the tools we implemented to evaluate our approach.
The implementation can be found at \url{https://github.com/p-otto/Armadillo}, in the folders \texttt{armadillo} and \texttt{eth\_admin}, respectively.

\subsubsection{Armadillo}

\begin{figure}
	\centering
\includegraphics[width=0.9\textwidth]{fig/Armadillo.png}
\caption{Screenshot of the Armadillo web application.}
\label{fig:armadillo}
\end{figure}

As a way of joining process execution with our blockchain architecture, we developed a single page web application (see Figure \ref{fig:armadillo}) in JavaScript and HTML, using the Vue.js framework\footnote{\url{https://vuejs.org}}.
With the help of the Web3 Ethereum JavaScript API\footnote{\label{web3}\url{https://github.com/ethereum/web3.js/}}, we can trigger contract functions based on user actions in the applications user interface.
Armadillo also implements uploading and rendering BPMN diagrams using the bpmn.js toolkit\footnote{\url{https://bpmn.io/toolkit/bpmn-js/}}.
Building upon the BPMN viewer provided in the toolkit, we implemented handling of user clicks on BPMN \emph{Service Tasks}, which allowed us to use the viewer as a placeholder for a process engine .
A screencast of the full choreography can be found in our repository: \url{https://github.com/p-otto/Armadillo/blob/master/armadillo-screencast.mp4}.

\subsubsection{Eth\_Admin}

We developed a small command line application for managing users and roles via the access contract.
The application is written in JavaScript using the node.js framework\footnote{\url{https://nodejs.org/}}.
Internally, it calls functions on access contracts utilizing the Web3 Ethereum JavaScript API\textsuperscript{\ref{web3}}.
We decided to build the command line functionality with the help of the Vorpal framework\footnote{\url{http://vorpal.js.org/}}.
The \texttt{README} in our repository describes how to use the application.

\section{Discussion} \label{evaluation}

When evaluating blockchain implementations, the underlying system comes with new aspects that need to be considered.
First, computation and data storage on the blockchain needs to be paid for in \emph{transaction fees}.
In Ethereum, transaction fees are paid for in \emph{gas}.
It is a unit that is used up by any instruction a contract executes.
Users can set gas prices for transactions, which affect how likely it is that the transaction gets executed by miners.
Therefore, higher gas prices lead to faster execution of transactions.

We measured the gas usage for each transaction that is executed by the Armadillo tool, and we summed it up for the different kinds of contracts in the companies in Table \ref{gasusage}:

\begin{table}
	\centering
	\caption{Gas used by contracts in the example choreography.} \label{gasusage}
	\begin{tabular}{|l | c | c | c|}
		\hline                          & \textbf{Factory} & \textbf{Instance} & \textbf{Sum}       \\
		\hline \textbf{Bulk Buyer}      & 4,590,569        & 59,396            & 4,649,965          \\
		\hline \textbf{Manufacturer}    & 129,588          & 124,316           & 253,904            \\
		\hline \textbf{Middleman}       & 129,586          & 111,189           & 240,775            \\
		\hline \textbf{Supplier}        & 86,398           & 61,745            & 148,143            \\
		\hline \textbf{Special Carrier} & 86,376           & 81,945            & 168,321            \\
		\hline \textbf{Sum}             & 5,022,517        & 438,591           & \textbf{5,461,108} \\ \hline
	\end{tabular}
\end{table}

Because of fluctuating exchange rates of USD to Ether and Ether to gas, the total gas cost varies.
At the time of writing the total cas cost amounted to \$2.50 according to \url{https://ethgasstation.info}, based on a gas exchange rate of 1 gas = 1 Gwei.

Most of the transactions have proven to come with reasonable gas costs.
However, the instantiation cost that is currently paid by the initiator appears to be problematic.
As the first instantiation transitively triggers all the other participant's instantiations, this can become quite costly, and might even surpass the \emph{block gas limit}, the maximum amount of gas that can be spent per new block, depending on the number of participants involved.
%Even if that is not the case, a high gas cost is still a problem, since in Ethereum, by design, transactions with high gas costs are less likely to be included in a block by miners, possibly leading to long waiting times. % TODO why again?

Security is one of the main reasons to use blockchain systems.
In our case, we put effort into making the message exchange secure by adding multiple steps of access verification using the access contracts.
Access contracts themselves, however, only implement access verification by comparing the sender address with the address the contract was deployed with (the \emph{owner} address).
As the owner can give and remove access rights freely, another layer of off-chain security might make sense here to restrict the owner account from unwanted access.

\section{Future Work} \label{futurework}

We researched a system prototype for handling process choreographies on the blockchain.
While we were able to execute an example choreography using our implementation, a few aspects are left for future work.

\subsubsection{Instantiation Costs}

During evaluation, we saw that the instantiation costs are very high, have a potential to hit the gas limit in larger choreographies and also have to be paid fully by the initiator.
These problems need to be addressed before our solution is feasible for production use.

Further research thus could be invested into a more decentralized instantiation, where each participant pays for its own instance contract creation.
We propose to have the initiator emit a specific instance creation event that all other participants would listen to.
The process engine then reacts to such an event with the help of a trigger, and calls the \texttt{createInstance} method.
Then the other factory is notified that the instantiation was successful.

\subsubsection{Access Contract}

To implement access right management, we decided for a simple mapping-based approach, which requires large amounts of memory.
Storing large amounts data on the blockchain is expensive by design.
As the mappings grow in size with each user, role and task, costs rise.
Finding optimized and cheaper (both in computation cost and storage cost) approaches is part of future work.

An alternative approach for access control can be implemented based on pattern 5 of~\cite{xu2018pattern} - depicting real world permissions as tokens, and "paying" with these tokens on function call to verify ones role.

\subsubsection{Process Contract Generation}

Our contract implementations are the result of manual, iterative development.
Having reached the final iteration, we realized that the contracts share common patterns.

Most of the implementation can be generated, e.g. from a choreography diagram (or at least from the participants public processes).
The next step is to define a template for the factory contract source code, to determine which parts are easily generated, and which parts might need manual adjustments.
Given such a template, one can work on a contract generation tool that takes a diagram as input and produces factory contract code.
A possible implementation could use a rule-based approach for translating the choreography model, similar to~\cite{lopez2017caterpillar}.

\subsubsection{Process Engine Integration}

In related work~\cite{lopez2017caterpillar}, process logic was executed on the blockchain.
As we decided for a hybrid approach, using a process engine for execution and the Ethereum blockchain for traceable message passing, we implemented a placeholder process engine in our application.

The next step is to integrate the system with a productive process engine implementation, e.g. the Camunda BPMN Workflow Engine\footnote{\url{https://camunda.com/products/bpmn-engine}}, for further evaluation.

\section{Conclusion} \label{conclusion}

Currently, blockchain technology is most widely used to handle monetary transactions.
Instead of shifting assets between partners, however, we are using the blockchain to shift \emph{responsibilities}.
In this paper, we presented a system architecture and prototypical implementation for connecting process execution with blockchain technology.
While keeping the process execution off-chain, we developed an approach of inter-process communication using the Ethereum blockchain as a messaging channel.
Having a permanent ledger as the underlying data structure storing messages, faults and misunderstandings can be traced back to their origin and responsibilities are disclosed.
Building on the findings of this paper, improvements regarding runtime and execution costs open up a wide area of possible future work on the topic.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references}

\end{document}